---
title: "graphicsutils"
output:
  md_document
---


```{r echo=FALSE}
knitr::opts_chunk$set(
  fig.path = "inst/assets/img/",
  comment = "R>> ",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 5,
  fig.height = 5
)
```


# Description

The *graphicsutils* is an R package that adds various graphics utilities based on
the core package *graphics*. For now, the package is dedicated to store our
graphical functions and improve our coding skills. Note that similar functions
may already exist elsewhere (most likely in the [*plotrix*](http://cran.r-project.org/web/packages/plotrix/index.html) package).
However this package may help users to overcome some difficulties they may
encounter with *graphics*.

Also, as *graphicsutils*  is based on the *graphics* package, it
is not designed to work with the [*grid*](https://stat.ethz.ch/R-manual/R-devel/library/grid/html/grid-package.html) system (and thereby it does not work with [ggplot2](http://cran.r-project.org/web/packages/ggplot2/index.html)). Related
with this is one very helpful article by Paul Murrell:
[The gridGraphics Package](https://journal.r-project.org/archive/2015-1/murrell.pdf).



## Status

[![Build Status](https://travis-ci.org/inSileco/graphicsutils.svg?branch=master)](https://travis-ci.org/inSileco/graphicsutils)
[![Build status](https://ci.appveyor.com/api/projects/status/330p7f0djhpl998q?svg=true)](https://ci.appveyor.com/project/KevCaz/graphicsutils-qo99s)
[![codecov](https://codecov.io/gh/inSileco/graphicsutils/branch/master/graph/badge.svg)](https://codecov.io/gh/inSileco/graphicsutils)
[![](https://img.shields.io/badge/licence-GPLv3-8f10cb.svg)](http://www.gnu.org/licenses/gpl.html)

- Note that functions are written using Camel case (e.g. `showPalette()`)



# Installation

The easiest way to install `graphicsutils` is to use the
[*devtools*](http://cran.r-project.org/web/packages/devtools/index.html) package:

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("inSileco/graphicsutils")
```

Then, load it:

```{r, eval=FALSE}
library(graphicsutils)
```



# Main features

## Empty your plot

To start a figure from scratch it is often useful to get a plot without nothing
but having the correct size of axes. `plot0()` allows the user to do so:


```{r plot0}
plot0(c(0,1),c(0,1))
```

Quite empty, isn't it?
Also, it can be filled with any color using the `fill` parameter.


```{r plot0v2}
plot0(c(0,1), c(0,1), fill="lightskyblue1")
```


## Add a box

The `box2()` function allows the user to add any axes around the plot in a more flexible way.

```{r box2}
par(mar=rep(2,4))
plot0()
box2(which="figure", lwd=2, fill="grey30")
box2(side=12, lwd=2, fill="grey80")
axis(1)
axis(2)
```

## Encircle points

```{r encircle}
coords <- cbind(runif(10), runif(10))
plot0(coords)
points(coords, bg="grey25", pch=21)
encircle(coords, border="#7b11a1", lwd=2)
```

## Add an image

The `pchImage()` function eases the uses of `rasterImage()` to add images
(including png and jpeg files) on a graph. It allows to change the color of the whole image.

```{r pchImage}
pathLogo <- system.file("img", "Rlogo.png", package="png")
par(mar=c(4,1,4,1), mfrow=c(1,2))
plot0()
pchImage(0, 0, file=pathLogo, cex.x =4.5, cex.y=4)
plot0()
pchImage(0, 0, file=pathLogo, cex.x =4.5, cex.y=4, col="grey25", angle=25)
```


## A stacked areas chart

### A simple stacked areas


```{r envelop,  fig.width=6,  fig.height=6}
plot0(c(0, 10),c(0, 10))
sz <- 100
seqx <- seq(0, 10, length.out=sz)
seqy1 <- 0.2*seqx*runif(sz, 0, 1)
seqy2 <- 4+0.25*seqx*runif(sz, 0, 1)
seqy3 <- 8+0.25*seqx*runif(sz, 0, 1)
envelop(seqx, seqy1, seqy2, col="grey85", border=NA)
```

### A complete stacked areas


```{r stackedArea,  fig.width=6,  fig.height=6}
x <- data.frame(matrix(runif(200, 2, 10), 8, 25))
stackedAreas(x)
```

## Polar plot

```{r polarPlot,   fig.width=6,  fig.height=6}
polarPlot(1:40, stats::runif(40), to=1.9*pi, col="grey30", border="grey80")
```

## Get pretty ranges

```{r prettyRange}
vec <- stats::runif(20)
range(vec)
prettyRange(vec)
prettyRange(c(3.85, 3.88245))
```

## Interactive functions

Some functions are interactive and fairly understandable!
So, I suggest you try the following functions:

```{r eval=FALSE}
pickColors()
```

```{r eval=FALSE}
layout2()
```

## Colors

`darken()` and `lighten()` functions are convenient way to produce consistent
set of shaded colors with minimal effort; also use
`showPalette()` to display your palette.


```{r darken}
someblue <- darken("blue", 10*1:9)
showPalette(someblue)
```

```{r lighten}
somered <- lighten("red", 10*1:9)
showPalette(somered, add_codecolor=TRUE)
```

Since version 1.1-2 a set of color palettes has been added as a dataset.

```{r insileco}
utils::data(gpu)
showPalette(gpu$insileco, add_codecolor=TRUE)
```


# Access to Font-Awesome's icons

```{r getIcon}
names <- getIconNames()
my_icon <- getIcon(name='beer', col='grey80')
plotImage(my_icon)
```

# Vector fields

```{r vectorfields}
systLin <- function(X, beta){
    Y <- matrix(0,ncol=2)
    Y[1] <- beta[1,1]*X[1]+beta[1,2]*X[2]
    Y[2] <- beta[2,1]*X[1]+beta[2,2]*X[2]
    return(Y)
}
seqx <- seq(-2,2,0.31)
seqy <- seq(-2,2,0.31)
beta1 <- matrix(c(0,-1,1,0),2)
# Plot 1:
vecfield2d(coords=expand.grid(seqx, seqy), FUN=systLin, args=list(beta=beta1))
# Plot 2:
graphics::par(mar=c(2,2,2,2))
vecfield2d(coords=expand.grid(seqx, seqy), FUN=systLin,
   args=list(beta=beta1), cex.x=0.35, cex.arr=0.25,
   border=NA,cex.hh=1, cex.shr=0.6, col=8)
graphics::abline(v=0,h=0)
```


# To do list

- [ ] Improve the code coverage;
- [ ] Add a contributing section;
- [ ] Add a vignette gathering examples (I may create a simple cookbook website);
- [ ] add interactive mode in `showPalette()` function;
- [ ] `vectfield2d()` needs to be reviewed => I'll do so when I'll integrate nice looking arrows.
