---
title: "'graphicsutils': R functions to facilitate the use of 'graphics'"
output:
    md_document:
        variant: markdown_github
---


```{r echo=FALSE}
knitr::opts_chunk$set(
  fig.path = "inst/assets/img/",
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width=4,
  fig.height=4
)
```


# Description

*graphicsutils* is an R package that includes a set of graphical functions. As the [*plotrix*](http://cran.r-project.org/web/packages/plotrix/index.html), this package ass some graphics utilities based on the core package *graphics*. Note that this package is not intended to be sumitted to the CRAN for two reasons.
First, because many of these functions already exist in other packages (in a different form though).
Second, because I use it to improve my coding skill and my experience in writting a R package.
Nevertheless, this package may help users to deal with typical issues they may encounter.
If some of these functions turn out to be really valuable, then they will likely appear in a different package which will be submit to the CRAN. Also, *graphicsutils* is not intended to be used with [ggplot2](http://cran.r-project.org/web/packages/ggplot2/index.html) package. Note that

Travis: [![Travis](https://travis-ci.org/KevCaz/graphicsutils.svg?branch=master)](https://travis-ci.org/KevCaz/graphicsutils)



# Install

To install the package, use the [*devtools*](http://cran.r-project.org/web/packages/devtools/index.html) package.

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("KevCaz/graphicsutils")
```

Then, load it:

```{r}
library(graphicsutils)
```


# Main features

## Empty your plot

```{r}
plot0(c(0,1),c(0,1))
```
Empty isn't it ?

## Add a box

```{r}
par(mar=rep(2,4))
plot0()
box2(which="figure",lwd=2, col2fill="grey30")
box2(side=12, lwd=2, col2fill="grey80")
axis(1)
axis(2)
```

## A stacked areas chart

### A simple area


```{r,   fig.width=6,  fig.height=6}
plot0(c(0,10),c(0,10))
sz <- 100
seqx <- seq(0, 10, length.out=sz)
seqy1 <- 0.2*seqx*runif(sz, 0, 1)
seqy2 <- 4+0.25*seqx*runif(sz, 0, 1)
seqy3 <- 8+0.25*seqx*runif(sz, 0, 1)
envelop(seqx, seqy1, seqy2, col="grey85", border=NA)
```

### A complete stacked areas


```{r,   fig.width=6,  fig.height=6}
x <- data.frame(matrix(runif(200,2,10), 8, 25))
stackedAreas(x)
```

## Polar plot

```{r,   fig.width=6,  fig.height=6}
polarPlot(1:40, stats::runif(40), to=1.9*pi, col="grey30", border="grey80")
```

## Get pretty ranges

```{r}
vec <- stats::runif(20)
range(vec)
prettyRange(vec)
prettyRange(c(3.85,3.88245))
```

## Interactive functions

Some functions are interactive and fairly understandable !
So, I suggest you try:

```{r eval=FALSE}
pickColors()
```
and also

```{r eval=FALSE}
layout2()
```





# License
The *graphicsutils* package is licensed under the GPLv3 (http://www.gnu.org/licenses/gpl.html).



# To do list

1. Add more examples
2. Create a sustainable system to include different shapes (I am curenlty thinking about it).
2. grapdientPolygon must be completed to include images before exporting it.
2. Add interactive mode in 'showPalette Function'
2. Vectfield2d needs to be reviewed.
