---
title: "'graphicsutils' an overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Vignette Title}
abstract: "The package graphicsutils is a set of miscellaneous functions to be used with the graphics library."
author: Kevin Cazelles
---


# Introduction

The "graphicsutils" package proposes a set of functions that eases the use of
the `graphics` package. As this package is essentially a collection of functions,
this vignette is a guided tour of the package, a cook book that exemplifies
the functions implemented.

```{r, echo = FALSE}
library("graphicsutils")
```





# Empty your plot plot0()

It's often useful to empty a plot. The plot0 function makes it very easy to do so.
Moreover the argument `fill` color the plot region (it is equivalent to calling the `plotAreaColor()`).

```{r plot00, fig.height=6, fig.width=6, fig.cap="plot0() and plotAreaColor()"}
par(mar=c(1,1,1,1), mfrow=c(2,2))
#
plot0()
text(0, 0, "plot0")
#
plot0(fill="#8eb5e3")
#
plot0()
plotAreaColor(col="#8eb5e3")
#
plot0()
plotAreaColor(col="#8eb5e3", border="grey20", lty=2, lwd=4)
```

Also


```{r plot0}
plot0(c(0,1),c(0,1))
```

Quite empty, isn't it?
Also, it can be filled with any color using the `fill` parameter.


```{r plot0v2}
plot0(c(-10,10), fill='#ebebeb', grid.col = 'white')
```

And also add a text:

```{r plot0v3}
plot0(c(-10,10), asp=1, fill=8, text='cool', srt=45, cex=4, col=2)
```


# box2()

The `box2()` function allows the user to add any axes around the plot. A
limitation of the `box()` function is does not allowed all the combination of
axes. `box2()` proposes a simple way to select the sides to be added on the plot:
1 = bottom; 2 = left; 3 = top; 4 = right.

```{r box2, fig.height=6, fig.width=6}
par(mar=rep(2,4))
plot0()
box2(which="figure", lwd=2, fill="grey30")
box2(side=12, lwd=2, fill="grey80")
axis(1)
axis(2)
```


# circles()

```{r circles, fig.height=6, fig.width=6}
par(mar=c(1,1,1,1), mfrow=c(2,2))
plot0()
circles(0, 0, 1)
plot0()
circles(0, 0, seq(0,1,0.2))
plot0()
circles(0, 0, seq(0,1,0.2), lty=2, lwd=1.4)
plot0()
circles(0, 0 ,seq(0,1,0.2), from=pi, col="#BBBBFF88", lty=2, lwd=1.4)
```



# plotMeans()


```{r plotMeans, fig.height=3, fig.width=6}
dataset <- data.frame(dat=c(rnorm(50, 10, 2), rnorm(50, 20, 2)) , grp=rep(c('A','D'), each=50))
graphics::par(mfrow=c(1,3))
plotMeans(dat~grp, data=dataset, pch=19)
##
plotMeans(dat~grp, data=dataset, FUN_err= function(x) sd(x)*2, pch=15,
 ylim=c(-5,30), yaxs='i', connect=TRUE, args_con=list(lwd=2, lty=2, col='grey35'))
##
ser <- function(x) sd(x)/sqrt(length(x))
plot0(c(0,4), c(0,30))
plotMeans(dat~grp, data=dataset, FUN_err=ser, pch=15,
draw_axis=FALSE, add=TRUE, seqx=c(.5,3.5), mar=c(6,6,1,1), cex=1.4)
graphics::axis(2)
```


# polarPlot()

```{r, fig.height=6, fig.width=6}
polarPlot(1:40, stats::runif(40), to=1.9*pi, col='grey30', border='grey80')
```

# compassRose

```{r compassRose, fig.height=6, fig.width=6}
par(mfrow=c(2,2), mar=rep(1,4))
##
plot0(c(-1,1), asp=1)
compassRose(0, 0)
##
plot0(c(-1,1), asp=1)
compassRose(0.5, 0.5)
##
plot0(c(-1,1), asp=1)
compassRose(0, 0, cex.cr=2)
##
plot0(c(-1,1), asp=1)
compassRose(0, 0, rot=.75*pi, cex.cr=2, cex.let=1.5, offset=-1.25)
```

also

```{R compassRose2}
compassRose(0, rot=25, cex.cr = 2, col.let =2, add = FALSE)
```

# envelop

```{r envelop, fig.height=6, fig.width=6}
plot0(c(0,10),c(0,10))
sz <- 100
seqx <- seq(0, 10, length.out=sz)
seqy1 <- 0.2*seqx*runif(sz, 0, 1)
seqy2 <- 4+0.25*seqx*runif(sz, 0, 1)
seqy3 <- 8+0.25*seqx*runif(sz, 0, 1)
envelop(seqx, seqy1, seqy2, col='grey85', border=NA)
envelop(seqx, seqy2, seqy3, col='grey25', border=NA)
```



# Add images on plot


The `pchImage()` function eases the uses of `rasterImage()` to add images
(including png and jpeg files) on a graph. It allows to change the color of the whole image.

```{r pchImage}
pathLogo <- system.file("img", "Rlogo.png", package="png")
par(mar=c(4,1,4,1), mfrow=c(1,2))
plot0()
pchImage(0, 0, file=pathLogo, cex.x =4.5, cex.y=4)
plot0()
pchImage(0, 0, file=pathLogo, cex.x =4.5, cex.y=4, col="grey25", angle=25)
```

## plotImage

```{r plotImage2, fig.height=6, fig.width=6}
img <- png::readPNG(system.file('img', 'Rlogo.png', package='png'), native=TRUE)
op <- graphics::par(no.readonly = TRUE)
graphics::par(mfrow=c(2,2), mar=rep(1,4))
for (i in 1:4) plotImage(img)
graphics::par(op)
```

Another example:

```{r pchImage2, fig.height=6, fig.width=6}
img<-png::readPNG(system.file('img', 'Rlogo.png', package='png'), native=TRUE)
n<-15
plot0(c(0,1),c(0,1))
pchImage(0.1+0.8*stats::runif(n), 0.1+0.8*stats::runif(n), cex.x=0.2+1.6*stats::runif(n),
obj=img, angle=360*runif(n), col=2)
```


## Colors

## Creating shades of colors

The `darken()` and `lighten()` functions ease the creation of shades of a given
color. For any color palette, use the `showPalette()` to display them!


```{r darken, fig.height=6, fig.width=6}
someblue <- darken("blue", 10*1:9)
showPalette(someblue)
```

```{r lighten, fig.height=6, fig.width=6}
somered <- lighten("red", 10*1:9)
showPalette(somered)
```

```{r showpal, fig.height=6, fig.width=6}
showPalette(rainbow(81))
```

`darken()` and `lighten()` functions are convenient way to produce consistent
set of shaded colors with minimal effort; also use
`showPalette()` to display your palette.


```{r darken2}
someblue <- darken("blue", 10*1:9)
showPalette(someblue)
```

```{r lighten2}
somered <- lighten("red", 10*1:9)
showPalette(somered, add_codecolor=TRUE)
```

Since version 1.1-2 a set of color palettes has been added, see `gpuPalette()`.

```{r insileco}
showPalette(gpuPalette("insileco"), add_codecolor=TRUE)
```



## Encircle points

```{r encircle}
coords <- cbind(runif(10), runif(10))
plot0(coords)
points(coords, bg="grey25", pch=21)
encircle(coords, border="#7b11a1", lwd=2)
```

## Image 2

This function is designed for small matrices so the matrix is translated into
a colored image without any rotation of the axes.

```{r image2}
image2(matrix(1:9, 3))
image2(matrix(1:27, 3), from = 2, border = 2, lwd = 2)
```


# Stacked areas chart

## A simple stacked areas


```{r envelop2,  fig.width=6,  fig.height=6}
plot0(c(0, 10),c(0, 10))
sz <- 100
seqx <- seq(0, 10, length.out=sz)
seqy1 <- 0.2*seqx*runif(sz, 0, 1)
seqy2 <- 4+0.25*seqx*runif(sz, 0, 1)
seqy3 <- 8+0.25*seqx*runif(sz, 0, 1)
envelop(seqx, seqy1, seqy2, col="grey85", border=NA)
```

## A complete stacked areas

```{r stackedArea,  fig.width=6,  fig.height=6}
x <- data.frame(matrix(runif(200, 2, 10), 8, 25))
stackedAreas(x)
```


# Manipulation of figure dimensions

## Get pretty ranges

```{r prettyRange}
vec <- stats::runif(20)
range(vec)
prettyRange(vec)
prettyRange(c(3.849,3.88245))
```

## Howmany RC

This helper function returns a suggested number of row and column for a given
number of panles passed as an argument.

```{r howManyRC}
howManyRC(10)
howManyRC(16)
```




## Interactive functions

Some functions are interactive and fairly understandable!
So, I suggest you try the following functions:

```{r eval=FALSE}
pickColors()
```

```{r eval=FALSE}
layout2()
```

















### Polar plot

```{r polarPlot,   fig.width=6,  fig.height=6}
polarPlot(1:40, stats::runif(40), to=1.9*pi, col="grey30", border="grey80")
```




## Vector fields

```{r vectorfields,  fig.height = 6, fig.width = 6}
systLin <- function(X, beta){
    Y <- matrix(0,ncol=2)
    Y[1L] <- beta[1,1]*X[1L]+beta[1,2]*X[2L]
    Y[2L] <- beta[2,1]*X[1L]+beta[2,2]*X[2L]
    return(Y)
}
seqx <- seq(-2,2,0.31)
seqy <- seq(-2,2,0.31)
beta1 <- matrix(c(0,-1,1,0),2)
# Plot 1:
vecfield2d(coords=expand.grid(seqx, seqy), FUN=systLin, args=list(beta=beta1))
# # Plot 2:
graphics::par(mar=c(2,2,2,2))
vecfield2d(coords=expand.grid(seqx, seqy), FUN=systLin,
   args=list(beta=beta1), cex.x=0.35, cex.arr=0.25,
   border=NA,cex.hh=1, cex.shr=0.6, col=8)
graphics::abline(v=0, h=0)
```
